MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-09-05 chapters/MetaTracts.tex
updated image locations
shortened abstract, red thread in introduction

%REPLACED BY 
Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and future components. 
The materials of choice of a strongly growing number of industrial branches are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands.
Carbon fiber reinforced polymers, more specifically CFRPs with endless carbon fibers which we focus on in this work, typically consist of two main components: 
The matrix, which acts as a glue and the reinforcements, which allows for achieve the desired characteristics. 
Resins such as epoxy are typically used in a CFRP material system with endless carbon fibers for the matrix. 
For special applications additional filler materials, such as colors or other particles, may be integrated. 
However, the most important component is the reinforcement component, which consists in our application case of bundles of endless carbon fibers. 
The carbon fiber bundles are woven into sheets of carbon fiber cloth, which are layered and reinforced by an epoxy material to manufacture a CFRP laminate.
The alignment and weaving pattern of a CFRP laminate strongly influences the strength and stiffness properties of the material. The resulting material shows high specific stiffness, high specific strength, and high corrosion resistance. 
Increasingly, complex primary structures and highly loaded components are being manufactured from CFRP.
Advanced composites and in particular carbon fiber reinforced polymers have thus successfully been introduced in aeronautic and automotive applications. Besides commercial aerospace and automotive industry, CFRP is also being used
in a wide range of applications such as aeronautical, marine and civil engineering, automotive and wind turbine design,
and sporting goods manufacturing.

The increasing share of fiber reinforced polymers also generates a strong demand for non-destructive testing (NDT) techniques~\cite{Red2012} of such materials and components. Ultrasonic testing is the most commonly used method for non-destructive testing of fiber reinforce polymers.
It provides a quick, cost-efficient overview, but has low resolution and is imprecise. Recently, industrial 3D X-Ray computed tomography (XCT) has been applied to non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, taken throughout a full rotation of the specimen.

In XCT, the specimen is  placed on the rotary table between X-ray source and detector and penetrated by an incident X-rays of the source.
The X-rays pass through the specimen and are attenuated by the material in the specimen.
The detector transfers the X-rays in its scintillator layer into visible light. The light is then recorded in a 2D projection image. 
The table is rotated by a small amount and the the process is repeated~\cite{heinzl-2008-thesis}.

In state of the art devices, XCT can reach voxel sizes below $500 nm$,
it generates high resolution XCT volume data for comprehensive and detailed analyses of the test specimens. However, there is still a trade-off between viewport and image resolution. The magnification reached within a cone beam XCT scan is determined by the specified distances
between source and specimen as well as source and detector. The magnification directly influences both resolution and viewport: higher resolutions decreases the viewport or how much of the structure can be imaged while lower resolutions allow for larger viewports.
The domain experts are increasingly interested in visualizing the internal geometric structures present in the  weaving pattern of the fiber bundles themselves instead of high resolution studies of individual fibers.

We focus on datasets with larger viewports and lower resolutions where the single fibers are barely visible. 
Figure~\ref{fig:data-char} provides an example of our dataset. It also shows an example of the recurring bundle pattern called an \textit{unit cell}.

 This work is thus motivated from the recent progress in two interrelated fields: On the one hand, fiber reinforced polymer (FRP) components has gained wide application in a variety of industries  because of FRP's superior material and physical properties in comparison to conventional materials \cite{Karpat2012}. On the other hand the use of industrial 3D X-Ray computed tomography (XCT) based non destructive testing of fiber reinforced components has also gained popularity~\cite{Schilling2005}.

@@

\section {Experimental Results}
\label {sec:results}
%\begin{figure*}
%\centering
%\captionsetup[subfigure]{labelformat=empty}
%	%\subfloat[]{\includegraphics[width=0.10\textwidth]{imagesMT2014/crop-6-G.PNG}}
%	%\subfloat[]{\includegraphics[width=0.10\textwidth]{imagesMT2014/crop-6-H.PNG}}
%	%\subfloat[]{\includegraphics[width=0.35\textwidth]{imagesMT2014/crop-6-I.PNG}}
%	
%	\subfloat[(a) Dat set 2]{\includegraphics[width=0.7\linewidth]{imagesMT2014/prepreg/Capture8.PNG}}
%	\subfloat[(b) Data set 3]{\includegraphics[width=0.2\linewidth]{imagesMT2014/dataset203.png}}	
%	\caption{(a) Shows in order, the volume rendering of the second data set, a single (2D) slice, the orientation clustering bundles and the final bundles followed by proximity clustering. (b) Shows in order the third data set, a single (2D) slice and finally the extracted fiber bundles.}
%	\label{fig:crop-6}
%\end{figure*}
\begin{figure*}[t]
\centering
	\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
	\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
	\caption{(A) Orientation wise decomposition of data set Figure~\ref{fig:data-char}A into the extracted bundles and (B) all of them combined.(C) Shows the results of voxelization of B. (D) shows a 2D slice of the volume. (E,F) shows the extracted meshes.}
	\label{fig:crop-16-decomp}
\end{figure*}
\begin{figure}
\centering
	\includegraphics[width=0.35\textwidth]{imagesMT2014/crop-6.PNG}
	\caption{Data set with flat thin and compact bundles. A shows the volume rendering and a 2D slice with one of the boundaries marked in red, B shows the clusters according to individual orientation.C shows the complete result. D shows the voxelization of C.}
	\label{fig:prepreg}
\end{figure}

>>>>

%REPLACED BY 
Modern industry is increasingly demanding function orientation, integration, and efficiency of novel materials and future components. The material of choice of a strongly growing number of industrial branches are carbon fiber reinforced polymers (CFRP), which allow an integration of those continuously rising demands.

CFRP materials show characteristics as high specific stiffness, high specific strength, high corrosion resistance at low weight in comparison to conventional materials. At the same time using CFRPs highly complex and integrated components may be produced, which were previously impossible to manufacture.
Increasingly, primary structures and highly loaded components are being manufactured from CFRP.
Thus CFRPs have successfully been introduced in aeronautic and automotive applications. Besides aerospace and automotive industry, CFRP is also being used in a wide range of applications such as marine and civil engineering, automotive and wind turbine design, and also sporting goods manufacturing.

Carbon fiber reinforced polymer components, more specifically CFRP laminates with endless carbon fibers which we focus on in this work, typically consist of two main components: 
The matrix on the one side, which acts as a bonding component and the reinforcements on the other side, which allows for achieving the desired characteristics e.g. in terms of stiffness, strength or durability. 
To manufacture CFRP laminates with endless carbon fibers, various production processes are in industrial use. 
Most of these processes start with reinforcement component, weaving individual carbon fiber bundles into sheets of carbon fiber cloth in a predefined pattern. 
These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also refereed to as fabric.
Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth. 
Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the strength and stiffness properties. 
Next, resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component. 
The main function of this matrix component is to act as a bonding component between the individual carbon fiber bundles. 
For the matrix of CFRP material systems with endless carbon fibers typically resins such as epoxy are frequently used. 
In special applications additional filler materials, such as colors or other particles, may be integrated. 
After curing the resin the production process of the CFRP laminate is finished.

The increasing share of CFRPs and also the complexity of both material system as well as the final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}. 
Ultrasonic testing is yet the most commonly used method for non-destructive testing of carbon fiber reinforce polymers.
On the one hand ultrasonic testing provides a quick and cost-efficient overview. On the other hand it shows low resolution and may generate arbitrary results, e.g. due to the geometry of the final component.

Aside ultrasonic testing, industrial 3D X-Ray computed tomography (XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. 
XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, taken throughout a full rotation of the specimen. 
In industrial XCT, the specimen is placed on the rotary table between X-ray source and detector and penetrated by incident X-rays of the source.
The X-rays pass through the specimen and are attenuated by the material in the specimen.
The detector records 2D projection image depending on its principle, e.g. by transferring the X-rays in a scintillator layer into visible light.  
The specimen is rotated by a predefined angular step and the process is repeated~\cite{heinzl-2008-thesis}. 
In state of the art devices, XCT can reach voxel sizes below $500 nm$.
XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens. 
Unfortunately, there is still a trade-off between viewport and image resolution. 
The magnification reached within a cone beam XCT scan is determined by the specified distances
between source and specimen as well as source and detector. The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport showing more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.

In this work we focus on datasets with larger viewports and lower resolutions where the single fibers are not or barely visible. Our domain experts are mainly interested in visualizing the geometric structures in the weaving pattern of the fiber bundles themselves instead of high resolution studies of individual fibers.
Figure~\ref{fig:data-char} provides an example of our dataset. It also shows an example of the recurring fiber bundle pattern, called a \textit{unit cell}.
Our work is thus motivated from the recent progress in two interrelated fields: On the one hand, fiber reinforced polymer (FRP) components has gained wide application in a variety of industries  because of FRP's superior material and physical properties in comparison to conventional materials \cite{Karpat2012}. On the other hand the use of industrial 3D X-Ray computed tomography (XCT) based non destructive testing of fiber reinforced components has also gained popularity~\cite{Schilling2005}.

@@

\section {Experimental Results}
\label {sec:results}
%\begin{figure*}
%\centering
%\captionsetup[subfigure]{labelformat=empty}
%	%\subfloat[]{\includegraphics[width=0.10\textwidth]{imagesMT2014/crop-6-G.PNG}}
%	%\subfloat[]{\includegraphics[width=0.10\textwidth]{imagesMT2014/crop-6-H.PNG}}
%	%\subfloat[]{\includegraphics[width=0.35\textwidth]{imagesMT2014/crop-6-I.PNG}}
%	
%	\subfloat[(a) Dat set 2]{\includegraphics[width=0.7\linewidth]{imagesMT2014/prepreg/Capture8.PNG}}
%	\subfloat[(b) Data set 3]{\includegraphics[width=0.2\linewidth]{imagesMT2014/dataset203.png}}	
%	\caption{(a) Shows in order, the volume rendering of the second data set, a single (2D) slice, the orientation clustering bundles and the final bundles followed by proximity clustering. (b) Shows in order the third data set, a single (2D) slice and finally the extracted fiber bundles.}
%	\label{fig:crop-6}
%\end{figure*}
\begin{figure*}
\centering
	\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
	\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
	\caption{(A) Orientation wise decomposition of data set Figure~\ref{fig:data-char}A into the extracted bundles and (B) all of them combined.(C) Shows the results of voxelization of B. (D) shows a 2D slice of the volume. (E,F) shows the extracted meshes.}
	\label{fig:crop-16-decomp}
\end{figure*}
\begin{figure}
\centering
	\includegraphics[width=0.35\textwidth]{imagesMT2014/crop-6.PNG}
	\caption{Data set with flat thin and compact bundles. A shows the volume rendering and a 2D slice with one of the boundaries marked in red, B shows the clusters according to individual orientation.C shows the complete result. D shows the voxelization of C.}
	\label{fig:prepreg}
\end{figure}

### 2014-09-05 metaTracts-pvis2014.tex
updated image,
shortened abstract, red thread in introduction

%% Author and Affiliation (multiple authors with multiple affiliations)
\author{Arindam Bhattacharya\thanks{e-mail: arindamb86@gmail.com}\\ %
        \scriptsize The Ohio State University %
\and Christoph Heinzl\thanks{e-mail:christoph.heinzl@fh-wels.at}\\ %
     \scriptsize University of Applied Sciences Upper Austria. %
\and Rephael Wenger\thanks{e-mail:wenger@cse.ohio-state.edu}\\ %
    \scriptsize The Ohio State University}

@@

%% Abstract section.
\abstract{
Carbon fiber reinforced polymers are bundles of individual carbon fibers (henceforth referred to as fiber bundle) woven into sheets of carbon fiber cloth which are layered
and reinforced by an epoxy material. Industrial 3D X-ray computed tomography (XCT) generates a 3D volumetric representation of a scanned object, reconstructed from a series of 2D penetration images,
taken throughout a full rotation of the object.	We describe a technique for extracting and visualizing fiber bundles in fiber reinforced composites scanned with X-ray computed tomography. Current approaches focus on identifying and extracting individual fibers.
Our goal is to visualize the geometric structures of the weaving pattern by analyzing large  areas of  composite woven fabrics using XCT scans of low resolution. Individual fibers are barely visible in such scans. We apply and extend methods originally developed for diffusion tensor imaging (DTI) on the generated XCT data. In the first step a coarse version of integral curves is used to extract sections of the fiber bundles in the woven fabrics. We call these sections MetaTracts.
Next the extracted fiber bundle sections are clustered using a two step approach, wherein we first cluster according to orientation and then  cluster by proximity. We test our algorithm on a number of data sets and show that it effectively identifies and separates different fiber bundles. This work will provide a stepping stone to introduce novel non destructive testing techniques for reinforced fiber bundles analysis.
} % end of abstract

@@

\bibliographystyle{abbrv}
%%use following if all content of bibtex file should be shown
%\nocite{*}
\bibliography{xBundle}
\input {./chapters/supplementaryMaterial.tex}
\end{document}

>>>>

%% Author and Affiliation (multiple authors with multiple affiliations)
\author{Arindam Bhattacharya\thanks{e-mail: arindamb86@gmail.com}\\ %
        \scriptsize The Ohio State University %
\and Christoph Heinzl\thanks{e-mail:christoph.heinzl@fh-wels.at}\\ %
     \scriptsize University of Applied Sciences - Upper Austria%
\and Rephael Wenger\thanks{e-mail:wenger@cse.ohio-state.edu}\\ %
    \scriptsize The Ohio State University}

@@

%% Abstract section.
\abstract{
%Carbon fiber reinforced polymers (CFRP) typically consist of two main components: the matrix and the reinforcement. 
%Most important in order to achieve the desired characteristics of a CFRP is the reinforcement component. 
%In case of CFRPs the reinforcements typically consist of bundles of endless carbon fibers woven into sheets of carbon fiber %cloth, which are then layered and combined to a CFRP by the matrix component. 
%For quality control of CFRP materials and components industrial 3D X-ray computed tomography (XCT) is increasingly applied %as it allows detailed characterizations of the material in a non-destructive manner.

This work introduces MetaTracts, a novel method for extracting and visualizing individual fiber bundles as well as their weaving pattern in endless carbon fiber reinforced polymer (CFRP) materials, scanned by X-ray computed tomography (XCT).

For the purpose of characterizing larger areas of CFRP materials such as complete unit cells integrating the recurring weaving pattern, the proposed workflow is designed to handle XCT scans of low resolution, in which the individual fibers of the bundles to be extracted are not or just barely visible.  
The proposed workflow applies and extends methods originally presented for the field of diffusion tensor imaging in order to cope with the low data quality: 
First, a coarse version of integral curves is used to trace subsections of the individual fiber bundles in the woven CFRP materials. We call these sections MetaTracts.
In the second step these extracted fiber bundle sections (MetaTracts) are clustered using a two step approach, wherein the first step clusters according to orientation and the second step cluster by proximity of the MetaTracts. 
Finally, the tool generates volumetric representations as well as surface models of the extracted fiber bundles to be exported for further analysis.

We evaluate the proposed workflow on a number of real world datasets and demonstrate that MetaTracts effectively and robustly identifies and separates different fiber bundles. 
%mention user study here?%
%This work will provide a stepping stone to introduce novel non destructive testing techniques for reinforced fiber bundles analysis.
} % end of abstract

@@

\bibliographystyle{abbrv}
%%use following if all content of bibtex file should be shown
%\nocite{*}
\bibliography{xBundle}
%\input {./chapters/supplementaryMaterial.tex}
\end{document}
